---
title: "Chapter 4 - Exploring Data with Graphs"
author: "Sarah Hedvig Dahl Nielsen"
date: "24/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### Packages used
```{r packages}
pacman::p_load(ggplot2)
```



### Data

The data is from the Companion website https://studysites.uk.sagepub.com/dsur/study/articles.htm

```{r loading data}
facebookData <- read.delim("FacebookNarcissism.dat", header = TRUE)
examData <- read.delim("Exam Anxiety.dat", header = TRUE)
festivalData <- read.delim("DownloadFestival.dat", header = TRUE)
```


### Plots

#### Exercise 1
In the first exercise we use the facebookData. It contains the following four variables


  - id: a number indicating from which participant the profile photo came

  - NPQC_R_Total: the total score on the narcissism questionnaire

  - Rating_Type: whether the rating was for coolness, glamour, fashion or attractiveness (stored as strings of text)

  - Rating: the rating given (on a scale from 1 to 5)


##### Exercise 1.1
The aim is to plot the relationship between narcissism (NPQC_R_Total) against the general profile ratings (Rating)

```{r exercise 1.1 (Facebook)}
ggplot(facebookData, aes(NPQC_R_Total, Rating)) + geom_point(aes(color = Rating_Type), position = "jitter")
```

##### Exercise 1.2
Plot a graph that shows the pattern in the data using only a line.

```{r exercise 1.2 (Facebook)}
ggplot(facebookData, aes(NPQC_R_Total, Rating)) + geom_smooth(method = lm, se = F)
```

##### Exercise 1.3
  - Plot different coloured lines for the different types of rating (cool, fashionable, attractive, glamorous)
  - Add a layer displaying the raw data as points

```{r exercise 1.3 (facebook)}
ggplot(facebookData, aes(NPQC_R_Total, Rating, color = Rating_Type)) + geom_smooth(method = lm, se = F) + geom_point()
```


#### Exercise 2

In the second exercise we use the exam anxiety data. It contains the following 5 variables

  - Code: a number indicating from which participant the scores came
  
  - Revise: the total hours spent revising
  
  - Exam: mark on the exam as a percentage
  
  - Anxiety: the score on the EAQ
  
  - Gender: Whether the participant was male or female
  
##### Exercise 2.1
The aim is to plot anxiety levels against exam performance and do a linear regression.

```{r Exercise 2.1 (exam anxiety)}
ggplot(examData, aes(Anxiety, Exam)) + geom_point() + geom_smooth(method = "lm", color = "Red", se = F) + labs(x = "Exam Anxiety", y = "Exam performance %")
```

##### Exercise 2.2

In this exercise we will build on the scatterplot from 'Exercise 2' to see whether gender affects the reaction to exam anxiety.

```{r Exercise 2.2 (exam anxiety)}
ggplot(examData, aes(Anxiety, Exam, colour = Gender)) + geom_point() + geom_smooth(method = "lm", aes(fill = Gender), alpha = 0.1) + labs(x = "Exam Anxiety", y = "Exam performance %")
```


#### Exercise 3
In exercise 3 we use the festival data

##### Exercise 3.1
Making a histogram

```{r exercise 3.1 (festival)}
ggplot(festivalData,aes(day1)) + geom_histogram(binwidth = 0.4) + labs(x = "Hygiene (Day 1 of Festival)", y = "Frequency")
```

##### Exercise 3.2

Making a boxplot
```{r Exercise 3.2 (festival)}
ggplot(festivalData, aes(gender,day1)) + geom_boxplot() + labs(x = "Gender", y = "Hygiene (Day 1 of Festival)")
```

##### Exercise 3.3
Finding the outlier
```{r Exercise 3.3 (festival)}
festivalData[order(festivalData$day1),]

festivalData <- festivalData[-c(810),]
```


##### Exercise 3.4 
Making a density plot

```{r Exercise 3.4 (festival)}
ggplot(festivalData, aes(day1)) + geom_density() + labs(x = "Hygiene (Day 1 of Festival)", y = "Density Estimate")
```

